---
title: "CYO Project Capstone Edx"
author: "Hildsley Noome"
date: "11 February 2019"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("devtools")
install.packages("tidyverse")
install.packages("caret")
install.packages("R.utils")
library(devtools)
install_github("vqv/ggbiplot")
install_github("eddelbuettel/rbenchmark")

library(ggbiplot)
library(tidyverse)
library(caret)
library(R.utils)
library(rbenchmark)

download.file(url = "https://archive.ics.uci.edu/ml/machine-learning-databases/00401/TCGA-PANCAN-HiSeq-801x20531.tar.gz",destfile = "Data/RNA_SeqData.tar.gz")  # Downloads the dataset

untar(tarfile = "Data/RNA_SeqData.tar.gz",exdir = "Data/") #Unzips the dataset as was downloaded from uci

myfile_data <- read_csv(file = "Data/TCGA-PANCAN-HiSeq-801x20531/data.csv") # reads the csv file from the data file in the folder, contains rna seq data
myfile_labels_rows <- read_csv(file = "Data/TCGA-PANCAN-HiSeq-801x20531/labels.csv") # Reads the csv file that contains the information regarding the types of samples

colnames(myfile_data)[1] <- "sample_number" # Changes the X1 column name to sample_number

colnames(myfile_labels_rows)[1] <- "sample_number" # Changes the X1 column name to sample_number, same as myfile_data for tidying data later

rna_seq_dat <- data.frame(myfile_labels_rows) %>% left_join(myfile_data, by = "sample_number")

rna_seq_dat$Class <- as.factor(x = rna_seq_dat$Class)


```

#Introduction

The human body is extraordinary in the sense that it exists out of billions of cells, where each cell maintains homeostasis with thousands of different pathways. The cells then should also be able to communicate to each other even though they could be in completely different parts of the body. This complex interaction and pathways should be maintained while the cells grow. 

Cells go through a cycle commonly refered to as mitosis. This cycle is important to ensure that cells can grow and replicate. During mitosis it is crucial that the DNA is replicated for subsequent division of a single cell into two completely new cells. Sometimes the pathways and enzymes responsible for DNA replication make mistakes, and even when other redundant enzymes are also present, DNA damage and mutations may occur and carried over to the new cell. 

These mutations or DNA damage may be in crucial area of the DNA which is responsible for growth and maintanence of cells. In this case these damages would lead to uncontrolled growth of cells and 
a tumour may exist. Tumours that continue to grow uncontrollable and spread to different parts of the body is commonly refered to as cancer. Cancer causes approximately 8.5 million deaths per year.

This collection of data is part of the RNA-Seq (HiSeq) PANCAN data set, it is a random extraction of gene expressions of patients having different types of tumor: BRCA, KIRC, COAD, LUAD and PRAD
Samples (instances) are stored row-wise. Variables (attributes) of each sample are RNA-Seq gene expression levels measured by illumina HiSeq platform.
Cancer Types:
BRCA = Breast Invasive carcinoma
COAD = Colon Adenocarcinoma
KIRC = Kidney Renal clear cell carcinoma
LUAD = Lung Adenocarcinoma
PRAD = Prostate adenocarcinoma



#Analysis



#Results



#Conclusion



```{r cars}
summary(cars)
```


```{r pressure, echo=FALSE}
plot(pressure)
```

